% -*- fill-column: 85; -*-
%!TEX root = ../dissertation.tex
\section{Related Work}
\label{s:related}

Chapter~\ref{sec:related} provides detailed related work; this section
addresses related work not covered there. Our previous work~\cite{hotos-ava}
proposed many key ideas in \model; this paper fleshes out and evaluates those
ideas.

\paragraphbe{FPGA virtualization}
has a long history~\cite{codezero,plessl2005zippy,
score,tartan06asplos,virtualRC,huang09fpgavirt,brant2012zuma,rcmw,
intermediate-fabrics}. Most prior work relies on hardware-specific features,
focuses on sharing in a single protection domain~\cite{amorphos}, or
virtualization primitives~\cite{cascade}. \model can be combined with any of
these techniques to virtualize FPGA accelerators.

\paragraphbe{Nooks}~\cite{nooks} uses kernel-level interposition
mechanisms that are similar in spirit to \Model. \Model's compiler generates
components that, like the wrappers and XPC in Nooks, provide transparent
control across address space and machine boundaries. Object tracking and
shadow copies in Nooks' \emph{NIM} are similar to the object tracking and
shadow buffers in \model.

\paragraphbe{RPC frameworks}~\cite{grpc,thrift,corba,Yang1996} provide an
interface description language (IDL) and tools to easily implement those
interfaces. Unlike the \compiler language, these IDLs do not capture all the
semantics of \emph{existing} C interfaces required to implement a \novtechabbrv
API remoting design. \Compiler also generates code for controlling remote
resources.

\paragraphbe{Program specification languages}~\cite{mssal} allow programmers
to specify properties of functions and their behavior, and are generally used
to check correctness, either statically (e.g., with model checking) or
dynamically (e.g., by inserting checks in the program). While such languages
allow (nearly) arbitrary predicates on programs, they are not designed to
provide semantic information to other tools, In addition, these languages are
not designed to specify how API calls are performed, and do not support
features like state tracking. \Speclang is optimized to allow easy and
specific descriptions of APIs and how calls should be performed.

\paragraphbe{Foreign Function Interface} tools allow one language to call
functions written in another, such as C.
Some~\cite{swig} make use of C headers, but require manual annotations in many
common cases. Unlike \model, language specific DSLs~\cite{cython,jna},
do not support marshalling data structures and encapsulate rather than export
the C API.
Cross-language serialization standards and frameworks~\cite{protobuf, msgpack,
ecma404}, only provide serialization for a set of primitives and supported
constructs. The user must write code to translate their data structure into
the language of the framework and provide their own transport for the
serialized data.