% !TeX root = ../dissertation.tex

In many parallel computing domains, compute density and
programmability~\cite{nvidia_cuda, stone2010opencl, gregory2014c++} have
made GPUs the clear choice for efficiency and performance~\cite{gpu_apps}.
Popular machine learning frameworks such as Caffe~\cite{jia2014caffe},
Tensorflow~\cite{abadi2016tensorflow}, Microsoft CNTK~\cite{yu2014cntk},
and Torch7~\cite{collobert2011torch7} rely on GPU acceleration heavily.
GPUs have made significant inroads in HPC as well: five of the top seven
supercomputers in the world are powered by GPUs~\cite{top500-Nov2018}.

Despite much prior research~\cite{rossbach16vee, kaveri16vee,
cc-numa-gpu-hpca15, abhishek-ispass16} on GPGPU virtualization, practical
options currently available to providers of virtual infrastructure all involve
bypassing the hypervisor. The most commonly adopted technique is to dedicate
GPUs to single VM instances via PCIe pass-through~\cite{AWS-gpu,gVirt},
thereby giving up the consolidation and fault tolerance benefits of
virtualization. More recently, industry players such as VMware, Dell and
BitFusion have introduced user-space API-remoting~\cite{bitfusion-whitepaper,
kim2012snucl, rCUDAnew, vmCUDA,rCUDA} based solutions as an alternative to
pass-through. API-remoting recovers the consolidation and encapsulation
benefits of virtualization but bypasses hypervisor interposition. The absence
of hypervisor interposition results in multiple disjoint resource managers
(the remote user-space API executor and the hypervisor) with no insight into
each others' decisions, thereby leading to poor decision making, and
priority-inversion problems~\cite{rossbach2011ptask}.

To recover hypervisor interposition while maintaining low-overhead, we explore
retrofitting GPGPU support into a virtual GPU device: We added support for
OpenCL to an implementation of the SVGA~\cite{dowty2009gpu} design in Xen, by
implementing the key missing component---a compiler for SVGA's TGSI virtual
ISA.
This effort helped us realize that because GPUs already support vendor-specific
virtual ISAs (vISAs), the additional vISA provides little benefit.
In fact, we found that it harms performance by necessitating a translation
layer that obscures the program's semantic information from the final
vendor-provided compiler.
Drawing on this lesson, we adapted \Trillium to take a more flexible approach
to ISA virtualization: eliding it entirely when the host GPU stack bundles a
compiler (most do), and using LLVM IR, when necessary, to provide a common
target for GPGPU drivers.

\Trillium represents an unexplored point in the GPGPU virtualization design space:
hypervisor-mediated API-remoting.
\Trillium is an existence proof of a viable alternative design that preserves desirable
virtualization properties such as consolidation, hypervisor interposition, isolation,
encapsulation, etc., without requiring full hardware virtualization.
% \Trillium interposes one of the lowest layers in the guest GPGPU stack: the Gallium3D pipe-driver.
% Typically, this results in a single OpenCL call being broken into multiple pipe-driver
% calls.
\Trillium outperforms a full virtualization system from the literature
by up to 14$\times$ (5.5$\times$ on average) and outperforms the para-virtual
SVGA-like design by as much as 7.3$\times$ (5.4$\times$ on average).
% \hyu{Should mention the slowdown compared with pure user-space API remoting?}

% This enables efficient interposition through a low-overhead communication channel to the hypervisor --- shared memory.
% Since the SVGA design uses a virtual ISA (vISA),

% It also requires that compilers that implement these
%translations be competitive with GPU vendor compilers.  The missing
%ISA must be replaced by something: we consider variants that treat the
%front end code as IR, refactoring compilation to the hypervisor, and a
%variant that uses LLVM IR as the virtual ISA.  The emerging Vulkan
%framework~\cite{Vulkanspec} suggests an ISA called SPIR-V as a replacement; we do not
%evaluate a variant that uses SPIR-V.

%We consider variants of the design that address the now absent TGSI
%At a minimum,
%induced by the need to translate a guest GPU ISA
%to a single hypervisor-supported virtual ISA, and subsequently re-translate it
% to the ISA of the system's physical GPU.

% Flexible interposition and strong isolation mechanisms
% are critical for device management: a virtualization layer's
% primary goal is to enable isolated sharing across VMs. However, virtualization
% of the GPU ISA serves only compatibility, and often does so redundantly
% as GPGPU frameworks like OpenCL and CUDA subsume compilers
% into the device driver. Recognizing this redundancy,
% Trillium elides GPU ISA virtualization, relegating the translation from GPU
% source code to physical GPU ISA to the hypervisor-resident driver. While the separation of
% concerns requires a compiler in host or hypervisor, it vastly reduces
% complexity, eliminates a redundant translation layer, and ensures that the
% GPU compiler has a high-fidelity view of the target hardware, restoring
% optimization opportunities sacrificed by a design that relies on multiple translations.
%Moreover, raising the level of abstraction for the GPU ABI from an ISA to
%source code is not as heretical as one might think: OpenCL drivers are generally
%expected to ship with an integrated source-level compiler.

% We evaluated a \Trillium prototype against systems %from the literature
% representative of full virtualization, API remoting, and VMware's SVGA.


% \noindent This paper make the following contributions:
% \begin{compactitem}
% \item We show that API-remoting does not have to be done entirely in user-space
%   and that it can be hypervisor-mediated with minimal loss of performance.
%   % \hyu{With tolerable loss? Compared with native run, this statement is not true. At least the impl\&measurement in this paper don't prove this point.}
% \item We implement GPGPU support for an SVGA-like design in the Xen
%   hypervisor, by completing a long-missing element---the TGSI compiler---%
%   in order to leverage OpenCL support provided by the Mesa/Gallium graphics
%   stack for Linux, via the Clover~\cite{GalliumCompute-web} project.
% \item We propose an improved design called \Trillium that
%   removes the necessity for the vISA defined by SVGA resulting in dramatic
%   performance improvements.
% \item We provide the first (to our knowledge) comprehensive empirical and
%   qualitative comparison of a wide range of fundamental virtualization
%   techniques from the literature.
% \end{compactitem}
